ISSSV1337 Group 5 Report

### 2022-08-03 Report/Group 5 report v2.Rmd
Authors, editing.

---
title: "Group 5 Report"
author: "ISSSV1337 Group 5"
date: "2022-08-03"
output:
  pdf_document: 
    toc: true
    highlight: pygments
    extra_dependencies: "subfig"
---

@@

<!--# Vi burde vel ha navnene våre, siden rapporten skal lastes opp på Inspera?-->

<!--# Legg til et sted: The Norwegian Control Committee for Fundraising (Innsamlingskontrollen) also has some very basic financial data and information available about selected charitable organisations.-->

@@

The project assigned to our group was Oslo Philanthropic Exchange (OPX), 
a startup working to develop solutions to assist individuals and 
organisations wanting to donate to charities and designing a socially 
responsible investment strategy. Their main goal is to increase cash 
flow from private groups and individuals to charitable purposes. Therefore, the donor's trust 
in the charities’ abilities to carry out their promises is of paramount significance.

@@

 <!-- Needs work -->
Additionally, by creating Key Performance Indicators based on these data, OPX can use 
them to benchmark the relationship between financial assets and emergency 
relief obligations.

@@

2. Creating KPI(s) is important for the organization’s performance 
management, to influence its objectives, and to foster growth.

@@

## Phase 1: How to harvest and analyse the data? {.smaller}

@@

We discussed creating a webcrawler that could collect data across
different websites, but quickly decided that this would be too
difficult, and that it would only work with selected organizations. The
suggested solution was to create a code that could read PDF documents
that OPX downloaded themselves. Our code would analyze and organize the
uploaded data from the PDF documents.

However, different organizations publish financial data in different
ways in their yearly reports -- some financial statements were included
as tables, some as pictures, etc. This led us to phase 2.

@@

```{r, echo=FALSE, fig.cap="Kreftforeningen 2021: The financial report saved as text in the PDF"}
# Illustration some text some pictures in financial statements
knitr::include_graphics("Images/lagret_tekst.png") # Kreftforeningen 2021
```

```{r, echo=FALSE, fig.cap="Regnskogfondet 2018: The financial report saved as an image in the PDF"}
# Illustration some text some pictures in financial statements
knitr::include_graphics("Images/lagret_bilde.png") # Regnskogfondet 2018
```

## Phase 2: Exploring solutions {.smaller}

At this stage, we wanted to explore alternatives to analyzing PDF
documents. We discussed how to do so, and concluded that we had three
options: 
- Financial statements in yearly report PDF documents. 
- Webscrape Proff.no for financial statements. 
- Webscrape the Brønnøysund registry.

We divided the task among ourselves to explore whether these solutions
were feasible or not. The first option of using PDF documents still had
the issue of financial statements not always being published as text.
Also, we quickly found that Proff.no disallowed webscraping on their
website. The Brønnøysund registry did store downloadable XML documents
of all the financial statements on their website, but we discovered that
access to these files cost about half a million NOK.

We started to look at workarounds for this problem. It would still be
possible to download Proff.no HTML websites manually for each
organization. And, we attempted to request a temporary access to the
Brønnøysund registry's XML documents.

The Brønnøysund registry and Proff.no had different divisions of
accounting items than the ones OPX wanted. We found that we needed to
create an item path from the webpage items to the OPX items, and we
attempted to do so in Excel at first, so that we later could code this
into R. While doing so, we realized that it would be practically
impossible to extract the correct sums for the OPX accounting items from
the Brønnøysund items. This was e.g. because OPX had split the operating
income into public and private contributions, while the Brønnøysund
registry and Proff.no had not.

```{r, echo=FALSE, fig.cap="Accounting items for Kreftforeningen 2020 and Diabetesforbunet 2020, gathered from the Brønnøysund registry", fig.align='center'}
# Illustration brns vs OPX accounting items
knitr::include_graphics("Images/Brns_items1.png")
```

```{r, echo=FALSE, fig.cap="Accounting items for Kreftforeningen 2020 and Diabetesforbunet 2020, gathered from OPX", fig.align='center'}
# Illustration brns vs OPX accounting items
knitr::include_graphics("Images/OPX_items1.png")
```

```{r, echo=FALSE, fig.cap="An attempt at marrying the two sets of accounting items", fig.align='center'}
# Illustration brns vs OPX accounting items
knitr::include_graphics("Images/Marriage.png")
```

When also discovering that not all organizations publish their financial
statements, and certainly not for every year, we decided to give up on
collecting financial data for OPX. This is when we moved on to the third
and final stage.

```{r, echo=FALSE, out.width="100%", fig.cap="Regnskogfondet stopped posting financial statement PDFs on their website in 2018"}
# Illustration only published for some years example
knitr::include_graphics("Images/Regnskogfondet_stops2018.png")
```

```{r, echo=FALSE, out.width="100%", fig.cap="Kreftforeningen started posting financial statement PDFs on their website in 2019"}
# Illustration only published for some years example
knitr::include_graphics("Images/Kreftforeningen_starts2019.png")
```

## Phase 3: Scrapping the scraping {.smaller}

The OPX case requested two tasks -- webscrape financial data, and create
KPIs. During phase three, we scrapped all ideas and progress from the
first two stages, and focused on creating KPIs.

@@

We wanted to use webscraping to collect the data, so that we could have
automated the data input process. As described, we explored different
potential solutions. We experimented on coding such that data could have
been imported from XML files. However, once we realized that the
financial data was of little use, this experimentation stopped.

As we entered the third phase of brainstorming, we still needed some
example data so that we could test our KPI codes. For this, we used the
financial data that OPX had already gathered from different
organizations and organized into useful accounting items.

We copied these financial statements manually into an Excel document.
Perhaps a "simple" solution, however our code requires OPX to import
both an Excel document themselves into shiny, as well as a HTML document
from Proff.no.

```{r, echo=FALSE, out.width="100%", fig.cap="The data format used to upload into shiny"}
# Bilde av den "endelige" excel-versjonen som brukes til å laste inn i shiny
knitr::include_graphics("Images/Excel_final.png")
```

@@

Our data was gathered from OPXs own web page, and from Proff.no. As described 
in the previous parts, we wanted to gather data from the yearly reports or by 
scraping the Brønnøyund registry or Proff.no, but found this to be close to 
impossible. The data used is the one that OPX has gathered themselves directly 
from the organizations. This data was also used to test that our codes in R 
were correct.

We have also used Proff.no to visualize the annual incomes and some KPIs. 
This is data that the organizations have reported themselves to Proff.no. 
Since there are no regulations regarding reporting data for NGOs in Norway, 
the organizations have not been consistent with the data reporting. This 
explains the holes in the presented data.

<!--# Fra Visualisations (nedenfor): The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.-->

<!--Legg inn bilder her-->

<!-- # Legger inn fire grafer som viser manglende data i Proffs årsregnskap-->

@@

# Findings

>>>>

---
title: "ISSSV1337 Group 5 Report"
author:    | 
    | Amalie Halle Christensen, Ida Kristine Garthe
    | Torbjørn Skinnemoen Ottersen, Henrik Waage Rui
    | Agnieszka Sadlowska
date: "2022-08-03"
output:
  pdf_document: 
    toc: true
    highlight: pygments
    extra_dependencies: "subfig"
---

@@

@@

The project assigned to our group was Oslo Philanthropic Exchange (OPX), 
a startup working to develop solutions to assist individuals and 
organisations wanting to donate to charities and designing a socially 
responsible investment strategy. Their main goal is to increase cash 
flow from private groups and individuals to charitable purposes. Therefore, the donors' trust in the charities’ abilities to carry out their promises is of paramount significance.

@@

Additionally, by creating Key Performance Indicators based on these data, OPX can use them to benchmark the relationship between financial assets and emergency relief obligations.

@@

2. Creating KPI(s) is important for the organisation’s performance 
management, to influence its objectives, and to foster growth.

@@

## Phase 1: How to harvest and analyse the data?

@@

We discussed creating web scraping code that could collect data across
different websites, but quickly decided that this would be too
difficult, and that it would only work with selected organizations. The
suggested solution was to create code that could read PDF documents
that OPX downloaded themselves. Our code would analyse and organise the
uploaded data from the PDF documents.

However, different organisations publish financial data in different
ways in their annual reports -- some financial statements were included
as tables, some as pictures, etc. This led us to phase 2.

@@

```{r, echo=FALSE, fig.cap="Norwegian Cancer Society 2021: The financial report saved as text in the PDF."}
# Illustration some text some pictures in financial statements
knitr::include_graphics("Images/lagret_tekst.png") # Kreftforeningen 2021
```

```{r, echo=FALSE, fig.cap="Rainforest Foundation 2018: The financial report saved as an image in the PDF."}
# Illustration some text some pictures in financial statements
knitr::include_graphics("Images/lagret_bilde.png") # Regnskogfondet 2018
```

## Phase 2: Exploring solutions

At this stage, we wanted to explore alternatives to analysing PDF
documents. We discussed how to do so, and concluded that we had three
options:

- Financial statements in anual report PDF documents. 
- Web scrape Proff.no for financial statements. 
- Web scrape the Brønnøysund registry.

We divided the task among ourselves to explore whether these solutions
were feasible or not. The first option of using PDF documents still had
the issue of financial statements not always being published as text.
Also, we quickly found that Proff.no disallowed web scraping on their
website. The Brønnøysund registry did offer API access to all the financial statements on their website, but we discovered that
access to these data costs about half a million NOK.

We started to look at workarounds for this problem. It would still be
possible to download Proff.no HTML web pages manually for each
organisation. We also received access to some XML files from Statistics Norway, but this would not be a viable solution for OPX in the long term.

The Brønnøysund registry and Proff.no had different divisions of
accounting items than the ones OPX wanted. We found that we needed to
create an item path from the webpage items to the OPX items, and we
attempted to do so in Excel at first, so that we later could code this
into R. While doing so, we realised that it would be practically
impossible to extract the correct sums for the OPX accounting items from
the Brønnøysund items. This was e.g. because OPX had split the operating
income into public and private contributions, while the Brønnøysund
registry and Proff.no had not.

```{r, echo=FALSE, fig.cap="Accounting items for the Norwegian Cancer Society 2020 and the Norwegian Diabetes Association 2020, gathered from the Brønnøysund registry.", fig.align='center'}
# Illustration brns vs OPX accounting items
knitr::include_graphics("Images/Brns_items1.png")
```

```{r, echo=FALSE, fig.cap="Accounting items for the Norwegian Cancer Society 2020 and the Norwegian Diabetes Association 2020, gathered from OPX.", fig.align='center'}
# Illustration brns vs OPX accounting items
knitr::include_graphics("Images/OPX_items1.png")
```

```{r, echo=FALSE, fig.cap="An attempt at marrying the two sets of accounting items.", fig.align='center'}
# Illustration brns vs OPX accounting items
knitr::include_graphics("Images/Marriage.png")
```

When we also discovered that not all organisations publish their financial
statements, and certainly not for every year, we decided to give up on
collecting financial data for OPX. This is when we moved on to the third
and final stage.

```{r, echo=FALSE, out.width="100%", fig.cap="The Rainforest Foundation stopped posting financial statement PDFs on their website in 2018."}
# Illustration only published for some years example
knitr::include_graphics("Images/Regnskogfondet_stops2018.png")
```

```{r, echo=FALSE, out.width="50%", fig.cap="The Norwegian Cancer Society started posting financial statement PDFs on their website in 2019."}
# Illustration only published for some years example
knitr::include_graphics("Images/Kreftforeningen_starts2019.png")
```

## Phase 3: Scrapping the scraping

The OPX case requested two tasks -- web scrape financial data, and create
KPIs. During phase three, we scrapped all ideas and progress from the
first two stages, and focused on creating KPIs.

@@

We wanted to use web scraping to collect the data, so that we could have
automated the data input process. As described, we explored different
potential solutions. We experimented on coding such that data could have
been imported from XML files. However, once we realised that the
financial data was of little use, this experimentation stopped.

As we entered the third phase of brainstorming, we still needed some
example data so that we could test our KPI codes. For this, we used the
financial data that OPX had already gathered from different
organisations and organised into useful accounting items.

We copied these financial statements manually into an Excel document.
Perhaps a 'simple' solution, however our code requires OPX to upload
both an Excel document, as well as a HTML document
from Proff.no, to the dashboard.

```{r, echo=FALSE, out.width="100%", fig.cap="The data format used to upload to the dashboard."}
# Bilde av den "endelige" excel-versjonen som brukes til å laste inn i shiny
knitr::include_graphics("Images/Excel_final.png")
```

@@

Our data was gathered from OPX's own web page, and from Proff.no. As described 
in the previous parts, we wanted to gather data from the annual reports or by 
scraping the Brønnøyund registry or Proff.no, but found this to be impossible. The data used is the one that OPX has gathered themselves directly 
from the organisations. This data was also used to test that our codes in R 
were correct.

We have also used Proff.no to visualise the annual incomes and some KPIs. 
This is data that the organisations have reported themselves to Proff.no. 
Since there are no regulations regarding reporting data for NGOs in Norway, 
the organisations have not been consistent with the data reporting. This may 
explain the lacunae in the presented data (see e.g. figures 9 and 10).

@@

\newpage
# Findings
